---
title: 'Analysis of the television data set'
author: Arthur Allignol
output:
  html_document:
    toc: true
    css: /data/Templates/GeneralStyle.css
    fig_caption: true
---


```{r setup, include = FALSE}
require(knitr)
require(xtable)
require(ggplot)
require(ggthemes)
opts_chunk$set(fig.path = "graphics/", size = "scriptsize")
```

We read the data using the following command
```{r read_data}
tele <- read.csv("/data/Ulm/Teaching/SS_2014/Intro_Biostat_Comp/Illustrations/Television/data/television.csv", 
                 stringsAsFactors = FALSE, na.strings = "*")
head(tele)
```

The aim of the analysis is to assess whether there exists a relation
between the number of people per television set and life
expectancy. 

We first note a couple of facts:

- `r tele$Country[which.max(tele$life)]` is the country with the highest
  life expectancy while `r tele$Country[which.min(tele$life)]` has the
  lowest;
  
- "Unsurprisingly", the `r tele$Country[which.min(tele$tv)]`
  have the lower number of habitants per television set and is equal
  to `r tele$tv[which.min(tele$tv)]`.

We will first explore the data graphically in
[Section 1][Graphical exploration] and perform a simple analysis in
[Section 2][Linear model].

## Graphical exploration

We first display a scatterplot of `life` and `tv`. A trend is clearly
to be seen but is non-linear.

```{r trend_simple, fig.height = 8, fig.width = 10, fig.cap = "Scatterplot of life expectancy and number of individuals per television set", out.width = "500px", fig.show = "hold", fig.pos = "h", fig.align = "center"}
ggplot(tele, aes(x = tv, y = life)) +
  geom_jitter(size =3) +
  geom_smooth(method = "lm") +
  theme_economist()
```

We first display a scatterplot of `life` and `tv`. A trend is clearly
to be seen but is non-linear.

As a second step, we log-transform the variable `tv`. The relation
between life expectancy and log televisions is in the next figure.
The transformation leads to an (almost) perfect linear
relation between these two variables. 
```{r log, fig.height = 8, fig.width = 10, fig.cap = "Scatterplot of life expectancy and log number of individuals per television set", out.width = "500px", fig.show = "hold", fig.pos = "h", fig.align = "center"}
tele$log_tv <- log(tele$tv)
ggplot(tele, aes(x = log_tv, y = life)) +
  geom_jitter(size =3) +
  geom_smooth(method = "lm") +
  theme_economist()
```


## Linear model

We fit a linear model with life expectancy as outcome and the
log-transformed number of individuals per television set.

```{r model}
fit_lm <- lm(life ~ log(tv), tele)
```
The results are displayed below
```{r tab_lm, results = "asis"}
out <- summary(fit_lm)$coefficients
out[, 1:3] <- round(out[, 1:3], 2)
out[, 4] <- format.pval(out[, 4], digits = 2,
                        eps = 10^(-3))

print(xtable(out),
      caption.placement = "top", type = "html")
```
We can check the model fit by plotting the residuals versus the fitted
values.
```{r res_vs_fit, fig.height = 8, fig.width = 10, fig.cap = "Residuals VS fitted", out.width = "500px", fig.show = "hold", fig.pos = "h", fig.align = "center"}
df <- fortify(fit_lm)
ggplot(df, aes(.fitted, .resid)) +
  geom_point(size = 2)  +
  geom_smooth(se=FALSE) +
  scale_x_continuous("Fitted Values") +
  scale_y_continuous("Residual") +
  labs(title = "Residuals vs fitted") +
  theme_economist()
```
